---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r, import libraries, echo=F, message=F, warning=F}
library(igraph)
library(SingleCellExperiment)
library(S4Vectors)
library(stringr)
library(DT)

library(dplyr)
library(tidyr)
library(mclust)
library(ggplot2)
library(RColorBrewer)
library(scater)
library(Rphenoannoy)
library(diffcyt)
library(edgeR)
set.seed(101100)
```


```{r, Set wd and load data, echo=F, message=F, warning=F}
#set working directory 
wd <-dirname(getwd())

data_folder <-(file.path(wd,"sce_objects","stroma","CAF"))
plot_folder <- file.path(wd,"plots")
#RAW
stroma.fibro <- readRDS(file=file.path(data_folder,"all_stroma_STROMA_RAW.rds"))
vessel.CAF <-readRDS(file=file.path(data_folder, "CAF_from_vessel.rds"))
#all.fibro <- readRDS(file=file=file.path(data_folder,"all_fibros_RAW.rds"))
#workingfile
#all.fibro <- readRDS(file=file.path(data_folder,"all_stroma_STROMA_RAW.rds"))
```

```{r, merge CAFs from vessel, echo=F, message=F, warning=F, eval=F}
colnames(colData(stroma.fibro))[!colnames(colData(stroma.fibro)) %in%colnames(colData(vessel.CAF))]
colnames(colData(vessel.CAF))[!colnames(colData(vessel.CAF)) %in%colnames(colData(stroma.fibro))]

cur_DF <- vessel.CAF %>% colData() %>% data.frame %>%select(-contains("vessel")) %>% DataFrame

all.equal(paste(cur_DF$ImageNumber, cur_DF$CellNumber), paste(vessel.CAF$ImageNumber, vessel.CAF$CellNumber))

colData(vessel.CAF) <- cur_DF
stroma.fibro$vessel <-NULL
all.fibro <-cbind(stroma.fibro, vessel.CAF)

saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_RAW.rds"))
saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile.rds"))
all.fibro <-readRDS(file=file.path(data_folder,"all_fibros_workingfile.rds"))

```

```{r, Define stroma markers, echo=F, warning=F, message=FALSE}
all.marker <-rownames(all.fibro)

bad.marker <- c("Iridium191","Iridium193","CD11b","Histone H3") 
good.marker <- all.marker[!all.marker %in% bad.marker]
fibro.marker <-c("FSP1" ,"aSMA", "CD146" ,"Cadherin-11","FAP","VCAM1",   "IDO" ,"Podoplanin","Carbonic Anhydrase IX", "CD73","p75 (CD271)","CD10" ,"Vimentin","Fibronectin" ,"PDGFR-b","CD34","CXCL-12","CCL21","Ki-67" ,"CXCL13"  )

stroma.marker <-c("FSP1" ,"aSMA", "CD146" ,   "Cadherin-11","FAP","VCAM1",   "IDO" ,"Podoplanin","Carbonic Anhydrase IX", "CD73","p75 (CD271)","CD10" ,"Vimentin","PNAd","Fibronectin","LYVE-1" ,"PDGFR-b","CD34","vWF","CXCL-12","CCL21","Ki-67" ,"CXCL13"  )
print(fibro.marker)
```

```{r, subset CAF all, echo=F, message=F, warning=F, eval=F}
#split cells by ImageNumber
n_cells_dr <- 100
cs <- split(seq_len(ncol(all.fibro)), all.fibro$Tma_ac)
length(unique(all.fibro$Tma_ac))
#sample 'n_cells_dr' per ImageNumber
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))
#sub.test <- sce[,cs]
all.fibro.sub <- all.fibro[,cs]

#calculate percentage of cells subsetting results in
p<-dim(assay(all.fibro.sub))[2]/dim(assay(all.fibro))[2]*100

#results in % of all cells
print(paste("Subsetting results in a total number of",dim(assay(
  all.fibro.sub))[2],"cells, which corresponds to",round(p,digits=2), "% of all cells of the original sce." ))

saveRDS(all.fibro.sub, file=file.path(data_folder, paste("all_fibro_sub.rds")))
```

```{r, load subset CAF all, message=FALSE, warning=FALSE, echo=FALSE}
all.fibro.sub <- readRDS(file=file.path(data_folder, paste("all_fibro_sub.rds")))
all.fibro.sub <- readRDS(file=file.path(data_folder, paste("final_fibro_sub.rds")))
```

```{r, calculate umap CAF all, warning=F, message=F, echo=F, eval=FALSE}
p <-c(10,25,50,75,100)
#p=50
for(i in p){
all.fibro <- runUMAP(all.fibro,
                 exprs_values = "c_counts_asinh",
                 name = paste0("UMAP_p", i),
                 #use_dimred="PCA_20",
                 n_neighbors = i)
saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile.rds"))

}
saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile.rds"))


for(i in p){
all.fibro <- runTSNE(all.fibro,
                 exprs_values = "c_counts_asinh",
                 name = paste0("tSNE_p", i),
                 #use_dimred="PCA_20",
                 perplexity = i)
saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile.rds"))
}
#saveRDS(all.fibro.sub, file=file.path(data_folder, paste("all_fibro_sub.rds")))

```


**UMAP with good markers**
```{r,plot umap CAF all good marker, fig.width=12, fig.height=12, echo=F, message=FALSE,warning=FALSE}

u.map <-c("UMAP_p10","UMAP_p25","UMAP_p50","UMAP_p75","UMAP_p100")

u.map="UMAP_p50"
i="UMAP_p50"
#i = "tSNE_p100"
for(i in u.map){
dat <-as.data.frame(reducedDims(all.fibro)[[i]])
dat$cell <- rownames(dat)
dat.counts <-as.data.frame(t((assay(all.fibro,"c_counts_asinh_scaled"))))
dat.counts$cell <- rownames(dat.counts)
dat.all <-merge(dat, dat.counts, by.x="cell", by.y="cell")

dat.all.long <-pivot_longer(dat.all, cols = colnames(dat.all)[colnames(dat.all) %in% good.marker], names_to = "target", values_to = "counts")


p <-dat.all.long %>%
  ggplot(aes(x=V1, y=V2, color=counts))+
  facet_wrap(~target, scales = "free", ncol = 6)+
  geom_point(alpha=0.5, size=0.2)+
  scale_color_gradientn(colours=rev(brewer.pal(11, 'Spectral')), name='Counts')+
  ggtitle('')+
  theme(strip.background = element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
         strip.text = element_text(size=8)) 

#ggsave(filename=file.path(plot_folder, paste0("",i,"CAFmarker",".png",sep="")), plot=p, width=8, height=6)
plot(p)
}

#CAFType
cluster <- "cell_type"
p <-plotReducedDim(all.fibro.sub, paste(i), colour_by=paste(cluster), point_size=0.5,)+labs(x="UMAP-1", y="UMAP-2", title="UMAP") +scale_color_d3("category20")# text_by = paste(cluster),text_colour ="black",
plot(p)
ggsave(plot=p, file=file.path(plot_folder, paste0("",i,"_CAF_types.png")))
#CAFSubType
cluster <- "cell_subtype"
p <-plotReducedDim(all.fibro.sub, paste(i), colour_by=paste(cluster), text_by = paste(cluster),text_colour ="red", point_size=0.5,)+labs(x="UMAP-1", y="UMAP-2", title="UMAP")+scale_color_d3("category20") 
plot(p)
ggsave(plot=p, file=file.path(plot_folder, paste0("",i,"_CAF_subtypes.png")))

#}
```
**tsne with good markers**
```{r,plot umap CAF all good marker, fig.width=12, fig.height=8, echo=F, message=FALSE,warning=FALSE}

t.sne <-c("tSNE_p10","tSNE_p25","tSNE_p50","tSNE_p75","tSNE_p100")

for(i in t.sne){
dat <-as.data.frame(reducedDims(all.fibro)[[i]])
dat$cell <- rownames(dat)
dat.counts <-as.data.frame(t((assay(all.fibro,"c_counts_asinh_scaled"))))
dat.counts$cell <- rownames(dat.counts)
dat.all <-merge(dat, dat.counts, by.x="cell", by.y="cell")

dat.all.long <-pivot_longer(dat.all, cols = colnames(dat.all)[colnames(dat.all) %in% fibro.marker], names_to = "target", values_to = "counts")


p <-dat.all.long %>%
  ggplot(aes(x=V1, y=V2, color=counts))+
  facet_wrap(~target, scales = "free", ncol = 6)+
  geom_point(alpha=0.5, size=0.2)+
  scale_color_gradientn(colours=rev(brewer.pal(11, 'Spectral')), name='Counts')+
  ggtitle('')+
  theme(strip.background = element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
         strip.text = element_text(size=8)) 

#ggsave(filename=file.path(plot_folder, paste0("tSNE_p",i,"CAFmarker",".png",sep="")), plot=p, width=16, height=10)
plot(p)
}

#CAFType
cluster <- "cell_type"
p <-plotReducedDim(all.fibro.sub, paste(i), colour_by=paste(cluster), text_by = paste(cluster),text_colour ="black", point_size=0.5,)+labs(x="tSNE-1", y="tSNE-2", title="tSNE") +scale_color_d3("category20")
plot(p)
ggsave(plot=p, file=file.path(plot_folder, paste0("tSNE_",i,"_CAF_types.png")))
#CAFSubType
cluster <- "cell_subtype"
p <-plotReducedDim(all.fibro.sub, paste(i), colour_by=paste(cluster), text_by = paste(cluster),text_colour ="red", point_size=0.5,)+labs(x="tSNE-1", y="tSNE-2", title="tSNE")+scale_color_d3("category20") 
plot(p)
ggsave(plot=p, file=file.path(plot_folder, paste0("tSNE_p",i,"_CAF_subtypes.png")))

}
```

```{r Clustering vessel stroma cells using Rphenoannoy, fig.width=25, fig.height=12, message=F, warning=F, echo=F,eval=F}
hmcol<-rev(brewer.pal(11,"RdBu"))
rdylbu <-rev(brewer.pal(11,"RdYlBu"))

#marker.clustering <- good.marker[!good.marker %in% "Pan Cytokeratin + Keratin Epithelial"]
#cellnb <- 250000
#subsample <-  colnames(sce)[sample(length(colnames(sce)), cellnb)]
#sce_sub <- sce[, colnames(sce) %in% subsample]

#set ks for clustering
#i <- 100
k <- c(20:40)
#i<-10
#i=20
for (i in k) {
  
  #all marker
    all.fibro$RPmembership <- factor(Rphenoannoy(data = t(assay(all.fibro[rownames(all.fibro) %in% good.marker,],"c_counts_asinh")), k = i)[[2]]$membership)
    cluster <- paste0("rp_fibro_all_k",i)
  colnames(colData(all.fibro))[which(names(colData(all.fibro)) == "RPmembership")] <- paste0(cluster)
  agg_sce <-aggregateAcrossCells(all.fibro, ids=all.fibro[[cluster]], average=TRUE, use_exprs_values="c_counts_asinh_scaled" )

 
    scater::plotHeatmap(agg_sce,
              features = good.marker,
              #features = all.marker,
              #features = marker.clustering,
              exprs_values = "c_counts_asinh_scaled",
              symmetric = FALSE,
              zlim=c(0,1),
              color=rdylbu,
              sortrowss=FALSE,
              show_colnames=TRUE,
              color_columns_by = c(cluster),
              main=paste0("Heatmap CAF, ",cluster))
    
  #save sce clustering
#saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile.rds"))


#fibro marker

   all.fibro$RPmembership <- factor(Rphenoannoy(data = t(assay(all.fibro[rownames(all.fibro) %in% fibro.marker,],"c_counts_asinh")), k = i)[[2]]$membership)
    cluster <- paste0("rp_fibro_fibro_k",i)
  colnames(colData(all.fibro))[which(names(colData(all.fibro)) == "RPmembership")] <- paste0(cluster)
  agg_sce <-aggregateAcrossCells(all.fibro, ids=all.fibro[[cluster]], average=TRUE, use_exprs_values="c_counts_asinh_scaled" )

 
    scater::plotHeatmap(agg_sce,
              features = fibro.marker,
              #features = all.marker,
              #features = marker.clustering,
              exprs_values = "c_counts_asinh_scaled",
              symmetric = FALSE,
              zlim=c(0,1),
              color=rdylbu,
              sortrowss=FALSE,
              show_colnames=TRUE,
              color_columns_by = c(cluster),
              main=paste0("Heatmap CAF, ",cluster))
    
  #save sce clustering

}
#all.fibro <-readRDS(file=file.path(data_folder,"all_fibros_workingfile.rds"))
saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile.rds"))


```

```{r}
df <- data.frame("CellID"=fibro.sce$CellID,
                 "CAFtype" =fibro.sce$cell_type,
                 "CAFsubtype"= fibro.sce$cell_subtype)

cur_DF <- as_tibble(colData(all.fibro.sub)) %>% left_join(df, by = "CellID") %>% DataFrame()
all.equal(paste(cur_DF$ImageNumber, cur_DF$CellNumber), paste(all.fibro.sub$ImageNumber, all.fibro.sub$CellNumber))

colData(all.fibro.sub) <- cur_DF
rownames(colData(all.fibro.sub)) <-all.fibro.sub$CellID
```

#FLOWSOM clustering

```{r, cluster CAFs using FLOWSOM, echo=F, eval=FALSE, warning=F, message=F, fig.width=15, fig.height=8}
fibro.marker.cluster <-c("SMA","FAP", "Cadherin-11", "Carbonic Anhydrase IX","Collagen I + Fibronectin",
                 #"VCAM1",
                 "Indoleamine 2- 3-dioxygenase (IDO)","Podoplanin","MMP11","CD73",
                 #"MMP9",
                 "CD10","Vimentin","CD248 / Endosialin",
                 #"LYVE-1",
                 "PDGFR-b","CD34","CXCL12","CCL21","Ki-67",
                 #"Caveolin-1",
                 "CD146","PNAd")



hmcol<-rev(brewer.pal(11,"RdBu"))
rdylbu <-rev(brewer.pal(11,"RdYlBu"))
assay(all.fibro, "exprs")<-assay(all.fibro, "c_counts_asinh")

#run FlowSOM
re2 <- CATALYST::cluster(all.fibro, features =fibro.marker, verbose = FALSE, maxK = 50)
#re2 <- CATALYST::cluster(all.fibro, features =fibro.marker, verbose = FALSE, maxK = 40)
fibro.test <- all.fibro
fibro.test2 <- all.fibro
library(CATALYST)
library(FlowSOM)
cl <-c(45:50)
#i <-5
for (i in cl){
  #i=35
  cluster <- paste0("som_",i)
  fibro.test2[[cluster]] <- as.factor(cluster_ids(re2, paste0("meta",i)))
  agg_sce <- aggregateAcrossCells(fibro.test2, ids=fibro.test2[[cluster]], average=TRUE, use_exprs_values="c_counts_asinh_scaled")
#plot Heatmap
  scater::plotHeatmap(agg_sce,
              #features = fibro.marker,
              features=c(fibro.marker),
              exprs_values = "c_counts_asinh_scaled",
              #symmetric = FALSE,
              zlim=c(-0,1),
              color=rdylbu,
              sortrowss=FALSE,
              show_colnames=TRUE,
              color_columns_by = c("ids"),
              main=paste0("Heatmap Fibros, ",cluster))
}

df <-data.frame("som_50"=fibro.test2$som_50, "CellID" =fibro.test2$CellID)
write.csv(df, file=file.path(data_folder, "CAF_cluster_som_50_df_cellID.csv"))

cur_DF <- as_tibble(colData(all.fibro)) %>% left_join(df, by = "CellID") %>% DataFrame()
all.equal(paste(cur_DF$ImageNumber, cur_DF$CellNumber), paste(all.fibro$ImageNumber, all.fibro$CellNumber))

colData(all.fibro) <- cur_DF
rownames(colData(all.fibro)) <-all.fibro$CellID
saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile_clustering.rds"))
```

**Heatmap fibroblast cluster after FLOWSOM clustering som_30**
```{r, assign CAF levels, fig.width=15, fig.height=8, echo=F, message=FALSE,warning=FALSE, eval=FALSE}
fibro.test2 <-readRDS(file=file.path(data_folder,"all_fibros_workingfile_clustering.rds"))
all.fibro <-readRDS(file=file.path(data_folder,"all_fibros_workingfile_clustering.rds"))
hmcol<-rev(brewer.pal(11,"RdBu"))
rdylbu <-rev(brewer.pal(11,"RdYlBu"))

is.vCAF <-c(23,14,41,39,34,38,33)
is.tpCAF.CD10 <-c(25,44,47,37)
is.tpCAF.CD73 <-c(50)
is.iCAF.CD34 <-c(45,46)
is.CAF.CXCL13 <- c(12,21)
is.IDO.CAF <-c(49)
is.CCL21.CAF <-31
is.mCAF.Col.Cdh <-c(42,43,48, 22,27,28)
is.hypox.tpCAF <-c(32)
is.hCAF <-c(7)
is.vim<-c(6,9,2,8,11,3,20,10)
is.other <- c(4,5,19,29)
is.fibro <- c(15,13,26,1)
is.SMA <- c(40,30,24,16,17,18, 35, 36)

all.fibro$CAFsubtype[all.fibro$som_50 %in% is.CAF.CXCL13] <-"CXCL13_CAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.vCAF] <-"vCAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.tpCAF.CD10] <-"CD10_CAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.tpCAF.CD73] <-"CD73_CAF"

all.fibro$CAFsubtype[all.fibro$som_50 %in% is.iCAF.CD34] <-"CD34_CAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.IDO.CAF] <-"IDO_CAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.hypox.tpCAF] <-"CA9_CD10_CAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.hCAF] <-"hypoxia"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.CCL21.CAF] <-"CCL21_CAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.vim] <-"Vimentin high"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.other] <-"other"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.mCAF.Col.Cdh] <-"FN_Cdh11_mCAF"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.fibro] <-"Fibroblast"
all.fibro$CAFsubtype[all.fibro$som_50 %in% is.SMA] <-"SMA_Fibro"

#caf type
all.fibro$CAFType[all.fibro$som_50 %in% is.CAF.CXCL13] <-"CXCL13_CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.vCAF] <-"CD146_CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.tpCAF.CD10] <-"CD10_CD73CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.tpCAF.CD73] <-"CD10_CD73CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.iCAF.CD34] <-"CD34_CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.IDO.CAF] <-"IDO_CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.hypox.tpCAF] <-"CA9_CD10_CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.hCAF] <-"hypoxia"
all.fibro$CAFType[all.fibro$som_50 %in% is.CCL21.CAF] <-"CCL21_CAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.vim] <-"Vimentin_high"
all.fibro$CAFType[all.fibro$som_50 %in% is.other] <-"other"
all.fibro$CAFType[all.fibro$som_50 %in% is.mCAF.Col.Cdh] <-"FN_Cdh11_mCAF"
all.fibro$CAFType[all.fibro$som_50 %in% is.fibro] <-"Fibroblast"
all.fibro$CAFType[all.fibro$som_50 %in% is.SMA] <-"SMA_Fibro"


agg_sce <- aggregateAcrossCells(all.fibro, ids=all.fibro$som_50, average=TRUE, use_exprs_values="c_counts_asinh_scaled")

 #plot Heatmap

pdf(file=file.path(plot_folder, "HM_som50_CAFclustering.pdf"), width=14, height=6)
scater::plotHeatmap(agg_sce,
              #features = good.marker,
              features=fibro.marker,
              exprs_values = "c_counts_asinh_scaled",
              symmetric = FALSE,
              zlim=c(-0,1),
              color=rdylbu,
              sortrowss=FALSE,
              show_colnames=TRUE,
              color_columns_by = c("CAFsubtype","CAFType"),
              main=paste0("Heatmap Fibros"))
dev.off()
agg_sce <- aggregateAcrossCells(all.fibro, ids=all.fibro$CAFsubtype, average=TRUE, use_exprs_values="c_counts_asinh_scaled")

 #plot Heatmap
scater::plotHeatmap(agg_sce,
              #features = good.marker,
              features=fibro.marker,
              exprs_values = "c_counts_asinh_scaled",
              symmetric = FALSE,
              zlim=c(-0,1),
              color=rdylbu,
              sortrowss=FALSE,
              show_colnames=TRUE,
              color_columns_by = c("CAFsubtype","CAFType"),
              main=paste0("Heatmap Fibros, ",cluster))

saveRDS(all.fibro, file=file.path(data_folder,"all_fibros_workingfile_clustering.rds"))




```


#complex heatmap coloured by CAF type
```{r}
#tableau 20 colours
library("scales")

palettes <- ggthemes_data[["tableau"]][["color-palettes"]][["regular"]]
for (palname in names(palettes)) {
  pal <- tableau_color_pal(palname)
  max_n <- attr(pal, "max_n")
  show_col(pal(max_n))
  title(main = palname)
}

agg_sce <-aggregateAcrossCells(fibro.final[, fibro.final$CAFType!="Vimentin_high"], ids=fibro.final[, fibro.final$CAFType!="Vimentin_high"]$CAFType, average=TRUE, use_exprs_values="c_counts_asinh_scaled" )

agg_sce <- agg_sce[rownames(agg_sce) %in% fibro.marker,]

hm.genes <- rownames(agg_sce)
hm.val <- assay(agg_sce,"c_counts_asinh_scaled")[hm.genes,]

#Subtype
n.cells <-as.data.frame(table(fibro.final[, fibro.final$CAFType!="Vimentin_high"]$CAFType))
n.cells$sqrt_freq <- sqrt(n.cells$Frequency)
colnames(n.cells) <-c("ids","Frequency")
col.pal = list(Cluster = viridis(length(unique(n.cells$ids))), Subtype= (palettes$`Tableau 20`)$value[1:11])

#no cluster colour
col.pal <-list(Subtype= (palettes$`Tableau 20`)$value[1:11])

names(col.pal$Subtype) <- unique(fibro.final[, fibro.final$CAFType!="Vimentin_high"]$CAFType)

col.pal

#cell type
p<-Heatmap(hm.val,
           name="Cluster",
           col=colorRampPalette(rev(brewer.pal(n = 7, name =
  "RdYlBu")))(100),
  clustering_method_rows = "ward.D2",
  clustering_method_columns = "ward.D2",
  top_annotation = HeatmapAnnotation(Frequency=anno_barplot(n.cells$sqrt_freq,height = unit(2, "cm")),
                                     col=col.pal,
                                     Subtype=unique(fibro.final[,
                                                              fibro.final$CAFType!="Vimentin_high"]$CAFType))) #,col=col.pal
draw(p)


#Save complex heatmap as pdf
pdf(file=file.path(plot_folder, paste("C_HM_CAF_Types_n_sqrt.pdf")), width=10, height=10)
draw(p)
dev.off()

```




```{r, subset CAF final, message=F, warning=F, echo=F}
#fibro.test2.PNAd <- fibro.test2[,fibro.test2$CAFtype=="HEV"]
#fibro.test2.other<- fibro.test2[,fibro.test2$CAFtype=="other"]

fibro.final <- all.fibro[,all.fibro$CAFsubtype!="other"]
all.fibro.other <- all.fibro[,all.fibro$CAFsubtype=="other"]



saveRDS(all.fibro.other, file=file.path(data_folder,"all_fibros_workingfile_clustering_OTHER.rds"))

saveRDS(fibro.final, file=file.path(data_folder,"all_fibros_workingfile_clustering_minusOTHER.rds"))

#saveRDS(fibro.final, file=file.path(data_folder,"FINAL_Fibros.rds"))
fibro.final <-readRDS(file=file.path(data_folder,"all_fibros_workingfile_clustering_minusOTHER.rds"))
```

```{r, plot UMAP ALL FIBROS types, message=F, warning=F, echo=F, eval=F}
#cluster <- "som_40"
#plotReducedDim(all.fibro.sub[, all.fibro.sub$som_40==15], "UMAP_p50", colour_by=paste(cluster), text_by = paste(cluster),text_colour ="red", point_size=0.5,)+labs(x="UMAP-1", y="UMAP-2", title="UMAP") 

#cluster <- "som_45"
#plotReducedDim(all.fibro.sub, "UMAP_p50", colour_by=paste(cluster), text_by = paste(cluster),text_colour ="red", point_size=0.5,)+labs(x="UMAP-1", y="UMAP-2", title="UMAP") 

#CAFType
cluster <- "CAFType"
plotReducedDim(all.fibro, "UMAP_p50", colour_by=paste(cluster), text_by = paste(cluster),text_colour ="red", point_size=0.5,)+labs(x="UMAP-1", y="UMAP-2", title="UMAP") 

#CAFSubType
red.dim <-reducedDimNames(all.fibro)
for( i in red.dim){
cluster <- "CAFsubtype"
p <-plotReducedDim(all.fibro, paste(i), colour_by=paste(cluster), text_by = paste(cluster),text_colour ="red", point_size=0.5,)+labs(x="UMAP-1", y="UMAP-2", title="UMAP") 
plot(p)
}
```

```{r, calculate umap CAF all, warning=F, message=F, echo=F, eval=FALSE}
p <-c(10,25,50,75,100)
#p=50
for(i in p){
fibro.final <- runUMAP(fibro.final,
                 exprs_values = "c_counts_asinh",
                 name = paste0("UMAP_p", i),
                 #use_dimred="PCA_20",
                 n_neighbors = i)
#saveRDS(fibro.final, file=file.path(data_folder,"all_fibros_workingfile_clustering_minusOTHER.rds"))

}
saveRDS(fibro.final, file=file.path(data_folder,"all_fibros_workingfile_clustering_minusOTHER.rds"))


for(i in p){
fibro.final <- runTSNE(fibro.final,
                 exprs_values = "c_counts_asinh",
                 name = paste0("tSNE_p", i),
                 #use_dimred="PCA_20",
                 perplexity = i)
saveRDS(fibro.final, file=file.path(data_folder, paste("final_fibro_sub.rds")))
}
saveRDS(fibro.final, file=file.path(data_folder, paste("final_fibro_sub.rds")))

```

**UMAP with good markers**
```{r,plot umap CAF all good marker, fig.width=12, fig.height=8, echo=F, message=FALSE,warning=FALSE}

plot_folder <- file.path(wd,"plots")
u.map <-c("UMAP_p10","UMAP_p25","UMAP_p50","UMAP_p75","UMAP_p100")

for( i in u.map){
dat <-as.data.frame(reducedDims(fibro.final)[[i]])
dat$cell <- rownames(dat)
dat.counts <-as.data.frame(t((assay(fibro.final,"c_counts_asinh_scaled"))))
dat.counts$cell <- rownames(dat.counts)
dat.all <-merge(dat, dat.counts, by.x="cell", by.y="cell")

dat.all.long <-pivot_longer(dat.all, cols = colnames(dat.all)[colnames(dat.all) %in% fibro.marker], names_to = "target", values_to = "counts")


p <-dat.all.long %>%
  ggplot(aes(x=V1, y=V2, color=counts))+
  facet_wrap(~target, scales = "free", ncol = 6)+
  geom_point(alpha=0.5, size=0.2)+
  scale_color_gradientn(colours=rev(brewer.pal(11, 'Spectral')), name='Counts')+
  ggtitle('')+
  theme(strip.background = element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
         strip.text = element_text(size=8)) 

ggsave(filename=file.path(plot_folder, paste(i,"_fibro_fibromarker.png",sep="")), plot=p, width=16, height=10)
plot(p)


#CAFType
cluster <- "CAFType"
p <-plotReducedDim(fibro.final, paste(i), colour_by=paste(cluster), text_colour ="red", point_size=0.5,)+labs(x="tSNE-1", y="tSNE-2", title="tSNE") +scale_fill_tableau("Tableau 20")
plot(p)
ggsave(p,file=file.path(plot_folder,paste0( i,"_fibro_","CAFtype.png")) )

#CAFSubType
cluster <- "CAFsubtype"
p <-plotReducedDim(fibro.final, paste(i), colour_by=paste(cluster), text_colour ="red", point_size=0.5,)+labs(x="tSNE-1", y="tSNE-2", title="tSNE") +scale_fill_tableau("Tableau 20")
plot(p)
ggsave(p,file=file.path(plot_folder,paste0( i,"_fibro_","CAFsubtype.png")) )

}
```
**tsne with good markers**
```{r,plot umap CAF all good marker, fig.width=12, fig.height=8, echo=F, message=FALSE,warning=FALSE}

t.sne <-c("tSNE_p10","tSNE_p25","tSNE_p50","tSNE_p75","tSNE_p100")

for(i in t.sne){
dat <-as.data.frame(reducedDims(fibro.final)[[i]])
dat$cell <- rownames(dat)
dat.counts <-as.data.frame(t((assay(fibro.final,"c_counts_asinh_scaled"))))
dat.counts$cell <- rownames(dat.counts)
dat.all <-merge(dat, dat.counts, by.x="cell", by.y="cell")

dat.all.long <-pivot_longer(dat.all, cols = colnames(dat.all)[colnames(dat.all) %in% fibro.marker], names_to = "target", values_to = "counts")


p <-dat.all.long %>%
  ggplot(aes(x=V1, y=V2, color=counts))+
  facet_wrap(~target, scales = "free", ncol = 5)+
  geom_point(alpha=0.5, size=0.2)+
  scale_color_gradientn(colours=rev(brewer.pal(11, 'Spectral')), name='Counts')+
  ggtitle('')+
  theme(strip.background = element_blank(),
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.background=element_blank(),
         strip.text = element_text(size=8)) 

ggsave(filename=file.path(plot_folder, paste(i,"_fibro_fibromarker.png",sep="")), plot=p, width=16, height=10)
plot(p)

#CAFType
cluster <- "CAFType"
p <-plotReducedDim(fibro.final, paste(i), colour_by=paste(cluster), text_colour ="red", point_size=0.5,)+labs(x="tSNE-1", y="tSNE-2", title="tSNE") +scale_fill_tableau("Tableau 20")
plot(p)
ggsave(p,file=file.path(plot_folder,paste0( i,"_fibro_","CAFType.png")) )

#CAFSubType
cluster <- "CAFsubtype"
p <-plotReducedDim(fibro.final, paste(i), colour_by=paste(cluster),  point_size=0.5,)+labs(x="tSNE-1", y="tSNE-2", title="tSNE") +scale_fill_tableau("Tableau 20")
plot(p)
ggsave(p,file=file.path(plot_folder,paste0( i,"_fibro_","CAFsubtype.png")) )

}
```

#Table of fibro subtype per patient
```{r}
df <- prop.table(table(fibro.final[, fibro.final$CAFsubtype !="Vimentin high"]$SampleID,fibro.final[, fibro.final$CAFsubtype !="Vimentin high"]$CAFsubtype),margin = 1) %>% data.frame
colnames(df) <-c("Patient","celltype","freq")
df
sum(df$freq)
library(ggthemes)

p <-df %>%
  ggplot(aes(x=freq, y=Patient, fill=celltype))+ geom_bar(stat="identity")+scale_fill_tableau("Tableau 20")+
  theme_bw()
plot(p)
ggsave(plot=p, file=file.path(plot_folder, "IMC_CAFsubtype_propBar.pdf"))

#numbers per patient
df <- table(fibro.final[, fibro.final$CAFsubtype !="Vimentin high"]$SampleID,fibro.final[, fibro.final$CAFsubtype !="Vimentin high"]$CAFsubtype) %>% data.frame
colnames(df) <-c("Patient","celltype","freq")
df
sum(df$freq)
library(ggthemes)

p <-df %>%
  ggplot(aes(x=freq, y=Patient, fill=celltype))+ geom_bar(stat="identity")+scale_fill_tableau("Tableau 20")+
  theme_bw()
plot(p)
ggsave(plot=p, file=file.path(plot_folder, "IMC_CAFsubtype_nBar.pdf"))

df <- df%>% pivot_wider(id_cols="Patient", names_from="celltype",values_from = "freq")
df$sum <- rowSums(df[2:13])
print(df)

print(colSums(df[2:13]) %>% data.frame)
```

#Table of fibro type per patient
```{r}
df <- prop.table(table(fibro.final[, fibro.final$CAFType !="Vimentin_high"]$SampleID,fibro.final[, fibro.final$CAFType !="Vimentin_high"]$CAFType),margin = 1) %>% data.frame
colnames(df) <-c("Patient","celltype","freq")
df
sum(df$freq)
library(ggthemes)

p <-df %>%
  ggplot(aes(x=freq, y=Patient, fill=celltype))+ geom_bar(stat="identity")+scale_fill_tableau("Tableau 20")+
  theme_bw()
plot(p)
ggsave(plot=p, file=file.path(plot_folder, "IMC_CAFtype_propBar.pdf"))

#numbers per patient
df <- table(fibro.final[, fibro.final$CAFType !="Vimentin_high"]$SampleID,fibro.final[, fibro.final$CAFType !="Vimentin_high"]$CAFType) %>% data.frame
colnames(df) <-c("Patient","celltype","freq")
df
sum(df$freq)
library(ggthemes)

p <-df %>%
  ggplot(aes(x=freq, y=Patient, fill=celltype))+ geom_bar(stat="identity")+scale_fill_tableau("Tableau 20")+
  theme_bw()
plot(p)
ggsave(plot=p, file=file.path(plot_folder, "IMC_CAFtype_nBar.pdf"))

df <- df%>% pivot_wider(id_cols="Patient", names_from="celltype",values_from = "freq")
df$sum <- rowSums(df[2:12])
print(df)

print(colSums(df[2:12]) %>% data.frame)
```


#clean
#Differential abundance analysis loop   
```{r  Differential Abundance Analysis Types over Metastasis clean, fig.width=6, fig.height=4, warning=F, message=F,echo=F, out.width="50%"}
#Patient Number must be numeric
library(tidyr)
library(ggthemes)
#groupId must be factor
fibro.final$TLS <- ifelse(fibro.final$Area_Description=="TLS", "TLS","no TLS")
colData(fibro.final)<-as.data.frame(colData(fibro.final)) %>% 
  group_by(RoiID) %>% mutate(NR=cur_group_id()) %>% DataFrame()
dat <-as_tibble(colData(fibro.final))
#if necessary: change group_id labels


subtype <-c("Gender","Relapse","Chemo","Dist.Met","LN.Met","Relapse")
subtype <- "TLS"
#define the naming subtype aka Celltype
dat$cell_type <- dat$CAFType
j <-"cell_type"

plot_list <- list()
for (i in subtype) {
  #for(k in unique(dat$DX.name)){    #uncomment this if you want to split your loop e.g. by tumour type
   # dat.l <-subset(dat, DX.name==k) #uncomment this if you want to split your loop e.g. by tumour type

dat.l <-dat
k="both tumour types" # comment this if you want to split by e.g. tumour type
dat.l$group_id <-as.factor(dat.l[[i]])
dat.l<-dat.l %>% drop_na(paste(i))

test <- dat.l %>% select(c("NR",j,"group_id"))

test_wide <- pivot_wider(test,id_cols=c("NR","group_id"),names_from = "cell_type", values_from ="cell_type", values_fn = list(cell_type=length),names_prefix = "")

test_wide
test_wide[is.na(test_wide)] <- 0

design <- createDesignMatrix(
  test_wide[,c("NR","group_id")], cols_design = c( "NR","group_id")
)

contrast <- createContrast(c(rep(0, ncol(design)-1),1))
data.frame(parameters = colnames(design), contrast)

test_wide_norm <- test_wide %>% select(-c("NR","group_id"))
#Normalize
norm_factors <- calcNormFactors(t(test_wide_norm), method = "TMM")
y <- DGEList(t(test_wide_norm), norm.factors = norm_factors)

y <- estimateDisp(y,design)

fit <- glmFit(y, design)
lrt <- glmLRT(fit, contrast = contrast)
top <- edgeR::topTags(lrt, n = Inf, adjust.method = "BH", 
        sort.by = "none")
t.op <-as.data.frame(top)
t.op <- tibble::rownames_to_column(t.op, "FibroTypes")

plot_list[[k]] <-
ggplot(t.op, aes(reorder(FibroTypes,logFC), logFC)) +
  geom_col(aes(fill=PValue<0.05)) +
  coord_flip() +
  scale_fill_tableau()+
  labs(x="Celltype", y="logFC",
       title=paste("Phenotypes over",i,k))+
  theme_bw()+
  theme(strip.background = element_blank(),
        panel.background=element_rect(fill='white', colour = 'black'),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.text.y = element_text(size=10))


#plot(p)


  #} #uncomment this if you want to split by e.g. tumour type
  gridExtra::grid.arrange(grobs = plot_list)
  
    #save individual pdf plots for each variable
  #pdf(file=file.path(plot_folder, paste0("DA_Celltype_n_over_",i,".pdf")))
  #gridExtra::grid.arrange(grobs = plot_list)
  #dev.off()
}

t.op %>% ggplot(aes(x=logCPM, y=logFC, color=FibroTypes))+geom_point()+scale_color_tableau()
t.op %>% ggplot(aes(x=logCPM, y=logFC, color=PValue<0.05))+geom_point()+scale_color_tableau()
```

## Kruskal-Wallis / Wilcoxon Fibro category ~ clinical data excluding neoadjuvant therapy
```{r  KW W Fibro category proportions WO neo, fig.width=12, fig.height=10, message=FALSE, warning=FALSE, echo=FALSE}
#df <- as.data.frame(colData(fibro.final))
caf.sce$TLS <- ifelse(caf.sce$Area_Description=="TLS", "TLS","no TLS")

df <- as.data.frame(colData(caf.sce))
df$cell_type <- df$CAFsubtype
#remove neoadj patients
categories <- c("TLS") 
#clinical.data <- data.frame("RoiID"=colData(fibro.final)$RoiID, "TLS"=colData(fibro.final)$TLS) %>% distinct
clinical.data <- data.frame("RoiID"=colData(caf.sce)$RoiID, "TLS"=colData(caf.sce)$TLS) %>% distinct

plot_list <- list()

for (i in (categories)) {

    t <- table(df$cell_type, df$RoiID)
    t <- prop.table(t, margin=2)
    t <- as.data.frame(t)
    colnames(t) <- c("Phenotype", "RoiID","freq")
    t<-merge(t, clinical.data, by.x="RoiID", by.y="RoiID")
    tdat <-t
    tdat$i <-as.factor(tdat[[i]])
    tdat <-tdat %>% drop_na(i)
    tdat$Phenotype <-droplevels(tdat$Phenotype)
  
  for (j in unique(t$Phenotype)){
    tdat<-t
    tdat[[i]] <-as.factor(tdat[[i]])
    tdat <-tdat %>% drop_na(i)
    tdat$i <-as.factor(tdat[[i]])
    tdat$Phenotype <-droplevels(tdat$Phenotype)

    tdat <-subset(tdat, Phenotype==j)

    if (length(unique(tdat[[i]])) > 2)
      {
    #calculate p values, add position for plot
    stat.test <- tdat%>%
     group_by(Phenotype) %>%
     dunn_test(freq~i,p.adjust.method = "bonferroni")
    stat.test<-stat.test %>% select(-.y., -statistic)
    stat.test <- stat.test %>% add_xy_position(x = i)
   
    #create plot list
    plot_list[[j]]<-
      ggplot(tdat, aes(x= .data[[i]], y = freq, colour=.data[[i]]))+
      geom_boxplot()+
      geom_point()+
      facet_wrap(~Phenotype, scales="free", ncol=4, strip.position="top")+
      scale_color_viridis_d()+
      theme_bw()+
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
      labs( y="Proportion")+
     theme(axis.text.x = element_text(angle=45, hjust=1), axis.ticks.x = element_blank(),
              legend.position = "none")+
      stat_pvalue_manual(stat.test, label = "p.adj.signif", tip.length = 0, hide.ns = T)
    }
    
    else {
if(is.element(i,colnames(clinical.data)) == TRUE){
    
      tdat<-t
      tdat[[i]] <-as.factor(tdat[[i]])
      tdat <-tdat %>% drop_na(i)
      tdat$Phenotype <-droplevels(tdat$Phenotype)

      tdat <-subset(tdat, Phenotype==j)
      detach("package:dplyr", unload=TRUE)
      library(dplyr)
      
      pvalues <- tdat %>% 
      group_by(Phenotype) %>% 
      summarise(p=wilcox.test(tdat$freq~tdat[[i]], paired=F)$p.value)
      tdat <- merge(tdat, pvalues, by.x = "Phenotype", by.y ="Phenotype", all.x = TRUE)
      tdat$p.wt <- paste0('p=',round(tdat$p, digits=3))

      #Plot list, all js together over i
      plot_list[[j]]<-
        ggplot(tdat, aes(x= .data[[i]], y = freq, colour=p<0.05))+
        geom_boxplot()+
        geom_point()+
        facet_wrap(~Phenotype+p.wt, scales="free", ncol=4, strip.position="top")+
        theme(axis.title.x=element_text("Phenotype"))+
      
        scale_colour_manual(values = setNames(c(palette("Tableau 10")[1:2]),c(F, T)))+
        theme_bw()+
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
        labs(x=paste(i), y="Proportion", fill="Area_Phenotype")+
        theme(axis.text.x = element_text(angle=45, hjust=1), axis.ticks.x = element_blank(),
              legend.position = "none")
    }  
 
    }
   }  

  #plot
  gridExtra::grid.arrange(grobs = plot_list, ncol=6) #, ncol=round(length(unique(t$Phenotype)))
  
  #save out in individual pdfs for each variable
  #pdf(file=file.path(plot_folder, paste0("KW_W_CAF_Proportions_WOneo",i,".pdf")), width=10, height=6)
 # gridExtra::grid.arrange(grobs = plot_list, ncol=round(length(unique(t$Phenotype))/2))
  #dev.off()
}
```

```{r}

df.caf <- colData(caf.sce)%>% data.frame

df.caf%>%
  group_by(CAFType)%>%
  summarise(mean=mean(Compartment),
            median=median(Compartment))

df.caf%>%
  group_by(CAFType)%>%
  summarise(mean=mean(Compartment))


df.caf%>%
  group_by(CAFsubtype)%>%
  summarise(mean=mean(Compartment))


p <- 
  ggplot(df.caf[df.caf$CAFType=="CD10_CD73CAF"|
                  df.caf$CAFType=="IDO_CAF"|
                  df.caf$CAFType=="FN_Cdh11_mCAF"|
                  df.caf$CAFType=="CD34_CAF",], aes(x = Compartment)) + 
  geom_density(aes(color = CAFType), position = "identity", bins = 30, alpha = 0.4)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  xlim(c(-100,50))+
  geom_vline(xintercept=0)

plot(p)

ggsave(p, file=file.path(wd, "plots","Distance_plot.pdf"))
```

#distance
```{r, Density plot cell types over distance}
#select cluster
all.sce[, all.sce$celltype=="CAF"]
sub.df <- dat.stroma %>% subset(Stroma_Phenotype=="tumour promoting CAF" | Stroma_Phenotype=="matrix CAF"  | Stroma_Phenotype=="inflammatory CAF"|  Stroma_Phenotype=="vessel CAF")

#calculate mean for h-line
cdat <- ddply(sub.df, "Stroma_Phenotype",summarise, distance.mean=median(Distance))
p <- 
  ggplot(sub.df, aes(x = Distance)) + 
  geom_density(aes(color = Stroma_Phenotype, fill = Stroma_Phenotype), position = "identity", bins = 30, alpha = 0.4)+
  scale_colour_viridis_d()+
  scale_fill_viridis_d()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  xlim(c(-100,50))

plot(p)
ggsave(plot=p, file=file.path(results_folder, paste("CAF_density_distance.pdf")))
```